---
title: "Pstat 115 HW2"
author: "Kalvin Goode"
date: "2018/10/20"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
```

(1)

(a)

We see that $E(A)=\frac{120}{10}=12=\frac{12}{1}=E(B).$ So we expect to have same average of incidence of cancer. However, we see that $Var(A)=\frac{120}{10^2}=1.2<12=\frac{12}{1^2}=Var(B)$. Therefore, we are more certain on A because it has lower variance.

(b)

```{r}
y_A=c(12, 9, 12, 14, 13, 13, 15, 8, 15, 6)
y_B=c(11, 11, 10, 9, 9, 8, 7, 10, 6, 8, 8, 9, 7)

cat("Posterior distribution for A is gamma(",120+sum(y_A),",",10+length(y_A),")")

cat("Posterior distribution for B is gamma(",12+sum(y_B),",",1+length(y_B),")")

cat("Posterior mean for A is",(120+sum(y_A))/(10+length(y_A)))

cat("Posterior mean for B is",(12+sum(y_B))/(1+length(y_B)))

cat("Posterior variance for A is",(120+sum(y_A))/(10+length(y_A))^2)

cat("Posterior variance for B is",(12+sum(y_B))/(1+length(y_B))^2)

cat("The 95% quantile-based credible intervals for theta_A is (",qgamma(0.025,120+sum(y_A),10+length(y_A)),",",qgamma(0.975,120+sum(y_A),10+length(y_A)),")")
cat("The 95% quantile-based credible intervals for theta_B is (",qgamma(0.025,12+sum(y_B),1+length(y_B)),",",qgamma(0.975,12+sum(y_B),1+length(y_B)),")")

```

(c)
```{r}
n_0=c(1:50)
a=12*n_0+sum(y_B)
b=n_0+length(y_B)
a/b
plot(a/b,xlab="n_0",ylab="Expectation",pch=16)
```

(d)

According to the 95% quantile-based credible intervals for $\theta_A$ and $\theta_B$, we see that these two groups are not independent and are related. Therefore, it does not make sense to have $p(\theta_A),\theta_B)=p(\theta_A)\times p(\theta_B)$


(2)

(a)

The first expert has mean=3/2000 and variance=3/$2000^2$, and second expert has mean=7/1000 and variance=7/$1000^2$. The first expert has more confidence because the vaiance is smaller than that of second expert.

(b)

```{r}
curve((dgamma(x,3,2000)+dgamma(x,7,1000))/2,0,0.02,xlab="Mixture Prior Distribution", ylab="Density")
```


(c)

Let $c_1$ and $c_2$ be constants. By conjugation, we have $c_1gamma(x;(3+8),(2000+1767))=Lgamma(x;3,2000)$ and $c_2gamma(x;(7+8),(1000+1767))=Lgamma(x;7,1000)$. We know that $gamma(x;a,b)=\frac{x^{a-1}}{\Gamma(x)b^a}exp(\frac{-x}{b})$. Thus, $c_1=\frac{11!\times 2000^3}{3767^{11}\times 3!}$ and $c_2=\frac{15!\times 1000^7}{2767^{15}\times 7!}$ And the posterior distribution up to a proportionality is $\frac{1}{2}(c_1gamma(x;11,3767)+c_2gamma(x;15,2767))$

(d)

We have $p(\lambda|y)=\frac{L(\lambda;y)p(\lambda)}{K}=\frac{c_1gamma(x;11,3767)+c_2gamma(x;15,2767)}{c_1+c_2}$. Notice that, from (c), we have $\frac{c_1}{c_1+c_2}=0.287$ and $\frac{c_1}{c_1+c_2}=0.712$. Thus, $p(\lambda|y)\sim 0.287gamma(11,3767)+0.712gamma(15,2767)$

(e)

```{r}

y<-function(x)
{
  fx<-0.28766*dgamma(x,11,3767)+0.71233*dgamma(x,15,2767)
}
curve(0.28766*dgamma(x,11,3767)+0.71233*dgamma(x,15,2767),0,0.01,xlab="x",ylab="Density")
abline(v=3/2000)
abline(v=7/1000)
MLE=optimize(y,c(0,0.01),maximum=TRUE)
text(3/2000, 0, labels='Expert A')
text(7/1000, 0, labels='Expert B')
text(MLE, 0, labels='MLE')
abline(v=MLE)
cat("The MLE is approximately") 
MLE
```

